Long-term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-23 article/article.tex
Commit 2024-06-23 17:12

The distribution of the number of CE events as a function of OIR is
shown in Figure\nobreakspace{}\ref{fig:ovir-distribution}. Apparently,
there is an inverse relationship between the frequency of CE events and
OIR with an exponential-like decline. This is expected, as the stronger
the CE events are, the rarer the conditions favoring the occurrence of
CE events. For the majority (over \(62\,\%\)) of the CE events the OIR
is below the long-term average of \(39.9\,\text{W}/\text{m}^2\), while
about \(8.1\,\%\) of the events correspond to OIR larger than
\(100\,\text{W}/\text{m}^2\) and up to the highest value of
\(412.4\,\text{W}/\text{m}^2\). This distribution is indicative of the
magnitude and the probability of the expected CE events over
Thessaloniki. Similar distribution of CE events, albeit with larger OIR
values, has been reported by \citet{Vamvakas2020}, for the city of
Patras. This site is located \(2.5^\circ\) south of Thessaloniki and is
exposed to air masses coming mainly from the eastern Mediterranean
resulting in different cloud patterns, that may affect the
characteristics and magnitude of the CE events.

@@

The duration of the CE events is variable and can last for several
minutes or even more than an hour. In order to study the characteristics
of these consecutive events, we grouped them into bins of increasing
duration in steps of one minute.

We have identified 28468 groups of CE in the whole period of study,
where the group with the longest duration lasted 140 minutes on 07~July
2013.

Figure\nobreakspace{}\ref{fig:ceg-duration-distribution}) shows the
frequency distribution of the CE events according to their duration. We
conclude that although some groups of events last for more than an hour
about \(80\,\%\) have duration of less than 5 minutes.

@@

The relation between the duration of the CE groups and the excess
irradiation of the groups has also been investigated
(Figure\nobreakspace{}\ref{fig:group-2d}). Evidently, evens of high
excess irradiation have small duration and vice versa. The vast majority
of grouped events are associated with small excess irradiation (e.g.
\(<5\,\text{kJ}/\text{m}^2\)) and small duration
(e.g.~\(<5\,\text{min}\)) while groups with strong excess irradiation
and long duration are rare. Similar results on this relation, have been
reported by \citet{Zhang2018}, on a study using a far higher sampling
rate than ours.

@@

\caption{Seasonal statistics of the number of ECE occurancies for each month, normalized to the maximum occurancies on March. The box represents the values of the low $25\,\%$ percentile to $75\,\%$ percentile, where the thick horizontal line inside is the mean, the vertical lines extend to the max imum and minimum vales, the dots are outlier values, and the rhombus is the mean.}\label{fig:relative-month-occurancies-ECE}
\end{figure}

@@

\hypertarget{discussion-and-conclusions}{%
\section{Discussion and conclusions}\label{discussion-and-conclusions}}

@@

The magnitude of the ECE events doesn't exceeds the values reported from
sites with more favourable conditions for the phenomenon. Although, the
climatological characteristic of the ECE events, showed that the most
active months are spread on half of the year (March -- June and
September and October). We found that CE conditions, can have a duration
of more than an hour in rare cases, with the bulk of the cases having a
duration under 5 minutes. Some of the characteristics of CE and ECE
events we analysed, have strong similarities with results by
\citet{Vamvakas2020}, for a city southern of Thessaloniki, with the
analogues differences on the intensity of the solar radiation.

>>>>

The distribution of the number of CE events as a function of OIR is
shown in Figure\nobreakspace{}\ref{fig:ovir-distribution}. Apparently,
there is an inverse relationship between the frequency of CE events and
OIR with an exponential-like decline. This is expected, as the stronger
the CE events are, the rarer the conditions favoring the occurrence of
CE events. For the majority (over \(62\,\%\)) of the CE events the OIR
is below the long-term average of \(39.9 \, \text{W} / \text{m}^2\),
while about \(8.1\,\%\) of the events correspond to OIR larger than
\(100\,\text{W}/\text{m}^2\) and up to the highest value of
\(412.4\,\text{W}/\text{m}^2\). This distribution is indicative of the
magnitude and the probability of the expected CE events over
Thessaloniki. Similar distribution of CE events, albeit with larger OIR
values, has been reported by \citet{Vamvakas2020}, for the city of
Patras. This site is located \(2.5^\circ\) south of Thessaloniki and is
exposed to air masses coming mainly from the eastern Mediterranean
resulting in different cloud patterns, that may affect the
characteristics and magnitude of the CE events.

@@

The duration of the CE events is variable and can last for several
minutes or even more than an hour. In order to study the characteristics
of these consecutive events, we grouped them into bins of increasing
duration in steps of one minute. We have identified 28468 groups of CE
in the whole period of study, where the group of the longest duration of
140 minutes occurred on 07~July 2013.
Figure\nobreakspace{}\ref{fig:ceg-duration-distribution} shows the
frequency distribution of the CE events according to their duration. We
conclude that although some groups of events last for more than an hour
about \(80\,\%\) have duration of less than 5 minutes.

@@

The relation between the duration and the mean OIR of the groups has
also been studied (Figure\nobreakspace{}\ref{fig:group-2d}). Evidently,
evens of high excess irradiation have small duration and vice versa. The
vast majority of grouped events are associated with small excess
irradiation (e.g.~\(<5\,\text{kJ}/\text{m}^2\)) and small duration
(e.g.~\(<5\,\text{min}\)) while groups with strong excess irradiation
and long duration are very rare. Similar results of this relation have
been reported by \citet{Zhang2018}, on a study using a far higher
sampling rate than ours.

@@

\caption{Seasonal variability of the number of ECE events normalized to the maximum occurring in March, in the form of a box and whisker plot. The box contains the data between the lower $25\,\%$ and the upper $75\,\%$ percentiles. The thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values.}\label{fig:relative-month-occurancies-ECE}
\end{figure}

@@

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

@@

The magnitude of the ECE events does not exceeds the values reported for
sites with more favourable conditions for the phenomenon. The
climatological characteristic of the ECE events, showed that the most
active months are spread on half of the year (March -- June and
September and October). We found that CE conditions, can have a duration
of more than an hour in rare cases, with the bulk of the cases having a
duration under 5 minutes. Some of the characteristics of CE and ECE
events we analysed, have strong similarities with results by
\citet{Vamvakas2020}, for a city southern of Thessaloniki, with the
analogues differences on the intensity of the solar radiation.
